---
title: "Homework 2 Report"
output: html_document
---


## Name

Nicholas Nasser

### Setup

Read in the data with `read_csv()` and store the data as an R object named `dataset`. Check the data to make sure all of the expected observations and variables are there.


```{r}
## Load the data and any necessary packages here.

library(tidyverse)
library(ggplot2)

data = read_csv("maacs.csv.gz")
```

## Part 1

We will first consider the relationship between FEV1 and age. In general, it is expected that as children get older (and hence, larger in size), their FEV1 values should get higher. 

Consider the statement "FEV1 values in children are higher in older children relative to younger children". 

Write a function in R that takes the `dataset` object as an argument and returns `TRUE` if the statement above is true for the dataset and `FALSE` otherwise. 

NOTE: In order to write this function, you will need to translate the statement above into something that can be checked with the data. There are many ways in which you can do that translation correctly and you only need to pick one way here. 

NOTE: For this part, do not use any plots.

```{r}
## Write your function here

fev1_increase_check = function(df) {
  correlation = cor(df$age, df$fev1, use = "complete.obs")
  return (correlation > 0)
}

fev1_increase_check(data)
```

I chose to create a function that takes in our Mouse Allergen and Asthma data and subsequently calculates the correlation between the FEV1 and Age columns. If the correlation coefficient is positive (indicating an increase in FEV1 levels as children age), then the function returns TRUE. Otherwise, it returns FALSE. Since the output above is TRUE, we can conclude that based on the data, the statement "FEV1 values in children are higher in older children relative to younger children" is true unless proven otherwise.

## Part 2

Fit a linear regression model with FEV1 as the outcome and age as a predictor.

How much does FEV1 change for a 1-year increase in the child's age? 

```{r}
## Add your code here

fev1_lm = lm(fev1 ~ age, data = data)
coef(fev1_lm)
```


Write your data analysis statement interpreting the regression model here:

Analysis Statement: Based on the model coefficients, for every 1-year increase in a child's age, we can expect the child's FEV1 to increase by approximately 0.171.

## Part 3

Develop **three** supporting premises derived from the data that support the statement you wrote in Part 2. These can be plots, other summary statistics, or model results. 

NOTE: 

* At least one supporting premise should use a plot.

* Do not use the code you write in Part 1 as a supporting premise

```{r}
## Add your code here

# Premise 1

ggplot(data = data, aes(x = age, y = fev1)) +
  geom_point(na.rm = TRUE) +
  geom_smooth(method = "lm", se = FALSE, color = "blue", na.rm = TRUE) +
  labs(title = "FEV1 vs Age in Children with Asthma",
       x = "Age (years)",
       y = "FEV1 (liters)") +
  theme_minimal()

# Premise 2

data$age_group <- cut(
  data$age,
  breaks = c(0, 5, 10, 15, 20),
  labels = c("0-5", "6-10", "11-15", "16-20"),
  right = TRUE
)

ggplot(data, aes(x = age_group, y = fev1)) +
  geom_boxplot(fill = "lightblue", na.rm = TRUE) +
  labs(
    title = "Distribution of FEV1 by Age Group",
    x = "Age Group (years)",                      
    y = "FEV1 (liters)"                           
  ) +
  theme_minimal()

# Premise 3

summary(fev1_lm)

```


Write the three supporting premise statements here:

1. The scatter plot with the fitted regression line displays a positive relationship between Age and FEV1 levels, indicating that FEV1 levels increase as children get older.

2. The box plot demonstrating levels of FEV1 in different age groups for children with asthma illustrate higher median FEV1 values for children who are older.

3. The summary of our linear model describing the relationship between Age and FEV1 displays an extremely small p-value for our slope coefficient of 0.171, showing us that the positive relationship is statistically significant.

## Part 4

For each of the supporting premises above, write a function that takes the `dataset` object as an argument and returns `TRUE` if the supporting premise statement above is true for the dataset and `FALSE` otherwise. 

For statements involving plots, instead of returning `TRUE` or `FALSE`, you function should do two things:

1. Produce the plot that is used in the statement

2. Produce a hypothetical version of the plot in the event that the statement is true. This can be done using simulated data or by simply hand drawing a plot.


```{r}
## Function for supporting premise statement 1

premise1_func = function(df) {
  
  # Plot used in statement
  plot_1 = ggplot(data = df, aes(x = age, y = fev1)) +
    geom_point(na.rm = TRUE) +
    geom_smooth(method = "lm", se = FALSE, color = "blue", na.rm = TRUE) +
    labs(title = "FEV1 vs Age in Children with Asthma",
         x = "Age (years)", 
         y = "FEV1 (liters)") +
    theme_minimal()
  print(plot_1)
  
  # Hypothetical plot
  
  hyp_lm1 = lm(fev1 ~ age, data = df)
  slope = coef(hyp_lm1)[2]
  X = rnorm(150, mean = mean(df$age, na.rm = TRUE), sd = sd(df$age, na.rm = TRUE))
  Y =  slope * X + rnorm(150, mean = mean(df$fev1, na.rm = TRUE), 
                               sd = sd(df$fev1, na.rm = TRUE) * 0.5) - 1.5
  sim_data = tibble(x = X, y = Y)
  plot_2 = ggplot(data = sim_data, aes(x = x, y = y)) +
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE, color = "blue",) +
    labs(title = "Hypothetical FEV1 vs Age in Children with Asthma",
       x = "Age (years)",
       y = "FEV1 (liters)") +
    theme_minimal()
  print(plot_2)
}
```


```{r}
## Function for supporting premise statement 2

premise2_func = function(df) {
  
  # Plot used in statement
  
  plot_1 = ggplot(data = df, aes(x = age_group, y = fev1)) +
    geom_boxplot(fill = "lightblue", na.rm = TRUE) +
    labs(
      title = "Distribution of FEV1 by Age Group",
      x = "Age Group (years)",                      
      y = "FEV1 (liters)") +
    theme_minimal()
  print(plot_1)

  # Hypothetical plot
  
  age_group = rep(c("0-5","6-10","11-15","16-20"), each = 20)
  age_group = factor(age_group, levels = c("0-5", "6-10", "11-15", "16-20"))
  
  sim_data <- data.frame(
    age_group = age_group,
    fev1 = c(rnorm(20, 0.8, 0.2), rnorm(20, 1.5, 0.2),
             rnorm(20, 2.0, 0.2), rnorm(20, 2.5, 0.2)))
  
  plot_2 = ggplot(sim_data, aes(x = age_group, y = fev1)) +
    geom_boxplot(fill = "pink") +
    labs(title = "Hypothetical Distribution of FEV1 by Age Group",
         x = "Age Group (years)", y = "FEV1 (liters)") +
    theme_minimal()
  print(plot_2)
}
```


```{r}
## Function for supporting premise statement 3

premise3_func = function(df) {
  lm_summary = summary(fev1_lm)
  slope = coef(lm_summary)[2]
  p_value = coef(lm_summary)[8]
  if (slope > 0 & p_value < 0.05) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}
```

Execute each of your function and show that the produce the expected output.

```{r}
premise1_func(data)
premise2_func(data)
premise3_func(data)
```

Each of these functions support the three supporting premises from question three, respectively. For the first two premises, I created functions that show the initial plots I created in question three using our real data and then created hypothetical plots rendered using simulated data. Both of these hypothetical plots also support the first two supporting premise statements. For the third statement, I created a function that checks the model summary of the linear model describing the relationship between FEV1 and Age, validating whether or not the slope coefficient is greater than zero and the p-value of the slope coefficient is smaller than 0.05. Both conditions were met, hence the function's output being TRUE.

## Part 5

Describe one alternative to the primary statement "FEV1 values in children are higher in older children relative to younger children".

```{r}
age_group = rep(c("0-5","6-10","11-15","16-20"), each = 20) %>%
  factor(., levels = c("0-5", "6-10", "11-15", "16-20"))
  
sim_data <- data.frame(
  age_group = age_group,
  fev1 = c(rnorm(20, 1.7, 0.2), rnorm(20, 1.65, 0.2),
           rnorm(20, 1.75, 0.2), rnorm(20, 1.7, 0.2)))
  
ggplot(sim_data, aes(x = age_group, y = fev1)) +
geom_boxplot(fill = "pink") +
labs(title = "Hypothetical Distribution of FEV1 by Age Group",
     x = "Age Group (years)", y = "FEV1 (liters)") +
theme_minimal()
```

One alternative to the primary statement "FEV1 values in children are higher in older children relative to younger children" is that "FEV1 values do not change as children age". This is represented above with simulated code, where in a hypothetical situation we observe FEV1 levels that are relatively constant among all children with asthma in the study, regardless of age.

Create a fault tree for the alternative outcome describing how the alternative outcome could be realized in the data even if the primary statement were true.

Your fault tree should be created as a separate image and does not need to be created in R. Upload the image of the fault tree to Canvas.









